
---
title: "Determining the relationships between reported GA/GAA and predicted placenta cell type proportion. Identifying the effects of PFAS on GAA."
author: "Cynthia Perez"

---

## Introduction

This analysis reproduces results related to the epigenetic clock using placental DNA methylation data. The code is structured to load pre-processed beta values, phenotype data, and log-transformed PFAS exposure data, and then conduct downstream statistical modeling.

## Load Required Libraries

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(planet) #installation via biorender 
library(qgcomp)
library(ggpubr)
library(car)
library(stringr)  
```

## Load Data

```{r load-data, message=FALSE}
# Load normalized beta values
load("2_BMIQ_Adjsuted.RData")
load("ALLBETAS.FunNormBMIQ.RData")

# Rename columns to match sample names
colnames(FunNorm.BMIQ.ALL) <- pd$Sample_Name

# Load PFAS data and phenotype info
load("log_PFAS_pheno.Rdata")
```

## Data Cleaning and Subsetting

```{r data-cleaning}
# Retain only samples present in both datasets
cols_to_keep <- intersect(colnames(FunNorm.BMIQ.ALL), rownames(log_pfas))
FunNorm.BMIQ.ALL <- FunNorm.BMIQ.ALL[, cols_to_keep]

# Subset phenotype and PFAS data accordingly
pheno <- pheno[cols_to_keep, ]
log_pfas <- log_pfas[cols_to_keep, ]
log_pfas <- log_pfas %>%
  rename_with(~ str_replace_all(.x, "-", "_"))   # "L-PFOA" -> "L_PFOA"

# Append PFAS values to phenotype data
pfas_vars <- c("L_PFOA","PFHxS","PFNA","PFDA","s_PFOS")
pheno <- cbind(pheno, log_pfas[, pfas_vars])
```

## The variable gwg_kg (gestational weight gain) was in a separate dataset. Adding it to pheno dataset. 

```{r adding-gwg_kg}
load("ViCTER_maternal_environment_2024-05-29.RData")
weight =  wide_pfas_modeling %>% 
          filter(id %in% pheno$participant_id) %>%
          select(id, gwg_kg) %>%
          rename(participant_id = id)


pheno= merge(pheno, weight, by = "participant_id")
rm(list = ls()[!ls() %in% c("pheno")])
```


## Notes

- The following files are not publicly available:
  - `2_BMIQ_Adjsuted.RData`
  - `ALLBETAS.FunNormBMIQ.RData`
  - `log_PFAS_pheno.Rdata`
  - `ViCTER_maternal_environment_2024-05-29.RData`
- Raw beta values are available from (GEO): GSE288358 

## predict gestational age using clocks from planet then extract residuals from regression analysis (gest_age_in_weeks_edd is reported GA)
These variables were already featured in pheno dataset. This code chunk is showing how GAA variables were generated. We proceeded with CPC_GAA based on correlation test and participants used to generate the CPC clock were similar to our study sample population. 

```{r generate-GAA,echo=FALSE}
pheno$RPC_Age = predictAge(FunNorm.BMIQ.ALL, type = "RPC")
pheno$CPC_Age = predictAge(FunNorm.BMIQ.ALL, type = "CPC")
pheno$RRPC_Age = predictAge(FunNorm.BMIQ.ALL, type = "RRPC")

pheno$RPC_GAA = summarY(lm(RPC_clock~gest_age_in_weeks_edd, pheno))$residuals
pheno$RRPC_GAA = summary(lm(RRPC_clock~gest_age_in_weeks_edd, pheno))$residuals
pheno$CPC_GAA= summary(lm(CPC_clock~gest_age_in_weeks_edd, pheno))$residuals

## determine the proper clock to use for analysis based on correlation value 
cor.test(pheno$gest_age_in_weeks_edd, pheno$CPC_clock)
cor.test(pheno$gest_age_in_weeks_edd, pheno$RPC_clock)
cor.test(pheno$gest_age_in_weeks_edd, pheno$RRPC_clock)
```

## Determinng the relationship between GAA/reported GA and 5 predicted cell type propotions'
***for loop of linear regression model where GAA/GA is the independent variable and a cell type is the dependent variable adjusted for the following covariates: mom age at birth,  mom education, sex of child/placenta,  mom gestational weight gain (gwg_kg), and official enrollment category (lean or overweight), reported GA (only when GAA is the independent variable)**
```{r cells-age, warning=FALSE}
cells <-  c( "Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")
age_levels <- c("CPC_GAA", "gest_age_in_weeks_edd")

results_list <- list()

for (age in age_levels) {

  out <- vector("list", length(cells))

  for (k in seq_along(cells)) {
    
    pf <- cells[k]
if (age == "gest_age_in_weeks_edd"){
  form <- reformulate(
      c(
        pf,
        "mom_age_at_birth",
        "mom_education",
        "official_enroll_category",
        "gwg_kg"
      ),
      response = age
    )
} else {
    form <- reformulate(
      c(
        pf,
        "mom_age_at_birth",
        "mom_education",
        "official_enroll_category",
        "gest_age_in_weeks_edd",
        "gwg_kg"
      ),
      response = age
    )
}
    

    model <- lm(form, data = pheno)
    coef_mat <- coef(summary(model))
    ci <- confint(model)[pf, ]

    out[[k]] <- data.frame(
      age       = age,
      cell_type      = pf,
      estimate  = coef_mat[pf, "Estimate"],
      conf.low  = ci[1],
      conf.high = ci[2],
      p.value   = coef_mat[pf, "Pr(>|t|)"],
      row.names = NULL
    )
  }

  results_list[[age]] <- bind_rows(out)
}

# Separate results
GAA_df <- results_list$CPC_GAA
GA_df <- results_list$gest_age_in_weeks_edd

GAA_df
GA_df

```


## The main effect: GAA ~ PFAS + covariates
**for loop of linear regression model where a PFAS chemical is the independent variable and CPC_GAA is the dependent variable adjusted for the following covariates: mom age at birth,  mom education, sex of child/placenta, reported gestational age,   mom gestational weight gain (gwg_kg), and official enrollment category (lean or overweight) ** 
```{r main-GAA-PFAS}
pfas_vars <- c("L_PFOA","PFHxS","PFNA","PFDA","s_PFOS")
out <- vector("list", length(pfas_vars))

for (k in seq_along(pfas_vars)) {
  pf   <- pfas_vars[k]
  form <- reformulate(
    c(
      pf, "mom_age_at_birth", "mom_education", "childs_sex",
      "official_enroll_category", "gest_age_in_weeks_edd", "gwg_kg"
    ),
    response = "CPC_GAA"
  )

  fit <- lm(form, data = pheno)

  ## get estimate + p-value
  est  <- coef(summary(fit))[pf, "Estimate"]
  pval <- coef(summary(fit))[pf, "Pr(>|t|)"]

  ## add CI (default 0.95)
  ci   <- confint(fit)[pf, ]     # two-column matrix: lower, upper

  out[[k]] <- data.frame(
    PFAS       = pf,
    estimate   = est,
    conf.low   = ci[1],
    conf.high  = ci[2],
    p.value    = pval,
    row.names  = NULL
  )
}

PFAS_cpc_lm <- do.call(rbind, out)

PFAS_cpc_lm
```

## Interaction: GAA ~ PFAS * sex + covariates 
**for loop of linear regression model where a PFAS chemical*sex is the independent variable and CPC_GAA is the dependent variable adjusted for the following covariates: mom age at birth,  mom education, sex of child/placenta, reported gestational age,  mom gestational weight gain (gwg_kg), and official enrollment category (lean or overweight) ** 
```{r int-GAA-PFAS}
pfas_vars <-  c("L_PFOA","PFHxS","PFNA","PFDA","s_PFOS")
out <- vector("list", length(pfas_vars))   # pre-allocate

for (k in seq_along(pfas_vars)) {
  pf <- pfas_vars[k]

  ## build CPC_GAA ~ PFAS * sex + covariates
  form <- reformulate(
    c(
      pf,                                      # main PFAS effect
      "mom_age_at_birth",
      "mom_education",
      "gest_age_in_weeks_edd",
      "childs_sex",
      "official_enroll_category", 
      "gwg_kg",
      paste0("childs_sex:", pf)                # interaction term
    ),
    response = "CPC_GAA"
  )

  fit <- lm(form, data = pheno)

  ## identify the interaction row 
  coef_mat <- coef(summary(fit))
  int_row  <- nrow(coef_mat)

  ## 95 % CI
  ci <- confint(fit)[int_row, ]

  out[[k]] <- data.frame(
    PFAS      = pf,
    estimate  = coef_mat[int_row, "Estimate"],
    conf.low  = ci[1],
    conf.high = ci[2],
    p.value   = coef_mat[int_row, "Pr(>|t|)"],
    row.names = NULL
  )
}

PFAS_cpc_int <- bind_rows(out)
PFAS_cpc_int

```
## Sex stratified: GAA ~ PFAS + covariates
```{r sex-GAA-PFAS}
pfas_vars <- c("L_PFOA", "PFHxS", "PFNA", "PFDA", "s_PFOS")
sex_levels <- c("Female", "Male")

results_list <- list()

for (sex in sex_levels) {
  dat_sex <- pheno %>% filter(childs_sex == sex)

  out <- vector("list", length(pfas_vars))

  for (k in seq_along(pfas_vars)) {
    pf <- pfas_vars[k]

    form <- reformulate(
      c(
        pf,
        "mom_age_at_birth",
        "mom_education",
        "official_enroll_category",
        "gest_age_in_weeks_edd",
        "gwg_kg"
      ),
      response = "CPC_GAA"
    )

    model <- lm(form, data = dat_sex)
    coef_mat <- coef(summary(model))
    ci <- confint(model)[pf, ]

    out[[k]] <- data.frame(
      sex       = sex,
      PFAS      = pf,
      estimate  = coef_mat[pf, "Estimate"],
      conf.low  = ci[1],
      conf.high = ci[2],
      p.value   = coef_mat[pf, "Pr(>|t|)"],
      row.names = NULL
    )
  }

  results_list[[sex]] <- bind_rows(out)
}

# Separate results
PFAS_cpc_lm_f <- results_list$Female
PFAS_cpc_lm_m <- results_list$Male

PFAS_cpc_lm_f
PFAS_cpc_lm_m

```

