---
title: "Reproducing Epigenetic Clock Results "
output: html_document
date: "2024-02-13"
---
```{r}
setwd("~/GitHub/Planet_Epigenetic_GA_Clocks")
library(tidyverse)
library(planet)
library(qgcomp)
library(ggpubr)
load("2_BMIQ_Adjsuted.RData")
load("ALLBETAS.FunNormBMIQ.RData")
library(r2excel)
colnames(FunNorm.BMIQ.ALL) = pd$Sample_Name

load("/Users/cynthiaperez/WGCNA_PFAS/log_PFAS_pheno.Rdata")
## Drop the two samples that weren't run (G-155 & G-157):
# Get the intersection of column names to keep
cols_to_keep <- intersect(colnames(FunNorm.BMIQ.ALL), rownames(log_pfas))

# Subset FunNorm.BMIQ.ALL to keep only the columns present in cols_to_keep
FunNorm.BMIQ.ALL <- FunNorm.BMIQ.ALL[, cols_to_keep]

pheno <- pheno[colnames(FunNorm.BMIQ.ALL),]
#rownames(log_pfas) == rownames(pheno) are now true after this next line 
log_pfas <- log_pfas[colnames(FunNorm.BMIQ.ALL),]

#add PFAS data columns to pheno change names 
pheno = cbind(pheno, log_pfas[,c("L-PFOA","PFHxS","PFNA","PFDA","s-PFOS")])
colnames(pheno)[50] = "PFOA"
colnames(pheno)[54] = "PFOS"
```




```{r}
t.test(pheno$gest_age_in_weeks_edd[which(pheno$childs_sex == "Female")], pheno$gest_age_in_weeks_edd[which(pheno$childs_sex == "Male")])
```
```{r}
t.test(pheno$birth_len_cm[which(pheno$childs_sex == "Female")], pheno$birth_len_cm[which(pheno$childs_sex == "Male")])
```

```{r}
t.test(pheno$birth_wt_kg[which(pheno$childs_sex == "Female")], pheno$birth_wt_kg[which(pheno$childs_sex == "Male")])
```

```{r}
t.test(pheno$CPC_lm_residuals[which(pheno$childs_sex == "Female")], pheno$CPC_lm_residuals[which(pheno$childs_sex == "Male")])
```

```{r}
hist(pheno$gest_age_in_weeks_edd)
```

```{r}
p <- ggboxplot(pheno, x = "childs_sex", y = "gest_age_in_weeks_edd",
          add = "jitter", xlab = "Sex of Infant", ylab = "Gestational Age")
p + stat_compare_means(method = "t.test", label.x = 1.4, label.y = 42)
```

```{r}
p <- ggboxplot(pheno, x = "childs_sex", y = "CPC_lm_residuals",
          add = "jitter", xlab = "Sex of Infant", ylab = "Gestational Age Acceleration")
p + stat_compare_means(method = "t.test", label.x = 1.4, label.y = 3)
```
```{r}

p <- ggboxplot(pheno, x = "childs_sex", y = "birth_len_cm",
          add = "jitter", xlab = "Sex of Infant", ylab = "Birth Length")
p + stat_compare_means(method = "t.test", label.x = 1.4, label.y = 55)
```
```{r}

p <- ggboxplot(pheno, x = "childs_sex", y = "birth_wt_kg",
          add = "jitter", xlab = "Sex of Infant", ylab = "Birth Weight")
p + stat_compare_means(method = "t.test", label.x = 1.4, label.y = 4.6)
```


#Predict Epigenetic Age 
```{r}
pheno$RPC_clock = predictAge(FunNorm.BMIQ.ALL, type = "RPC")
pheno$CPC_clock = predictAge(FunNorm.BMIQ.ALL, type = "CPC")
pheno$RRPC_clock = predictAge(FunNorm.BMIQ.ALL, type = "RRPC")
```

#Extract redisudals (GAA)
```{r}
pheno$RPC_lm_residuals = summary(lm(RPC_clock~gest_age_in_weeks_edd, pheno))$residuals
pheno$RRPC_lm_residuals = summary(lm(RRPC_clock~gest_age_in_weeks_edd, pheno))$residuals
pheno$CPC_lm_residuals= summary(lm(CPC_clock~gest_age_in_weeks_edd, pheno))$residuals
```


#GAA ~ PFAS 
**for loop of GAA (produced by CPC) ~ PFAS with covariates mom age at birth, mom education, mom BMI 36 weeks, sex, Mom BMI 10 weeks, and official enrollment category** 
```{r}
PFAS_cpc_lm = data.frame()

#pheno = pheno %>% filter(.$CPC_lm_residuals > -0.01)
for (i in c("PFOA","PFHxS","PFNA","PFDA","PFOS")) {
  model = lm(paste0("CPC_lm_residuals~", i,"+ mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category "), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate,  PFAS = i)
  PFAS_cpc_lm = rbind(PFAS_cpc_lm, df)
}

PFAS_cpc_lm
```

#GAA ~ PFAS mixture 
```{r}
GAA_mixture = qgcomp(CPC_lm_residuals ~ PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)

GAA_mixture = data.frame(estimate =  GAA_mixture[["psi"]], p.value = GAA_mixture[["pval"]][2])
```

#sheet 1 of xlsx
```{r}
filename = "GAA_tables.xlsx"
wb = createWorkbook(type = "xlsx")
sheet = createSheet(wb, "Sheet 1")
xlsx.addHeader(wb, sheet, value = "GAA ~ PFAS individual analysis", level = 2, color = "black")
xlsx.addLineBreak(sheet, 1)
covariates = "covariates: mom_age_at_birth, mom_education, MomBMI_36wks, childs_sex, MomBMI_10wks, official_enroll_category"
xlsx.addParagraph(wb, sheet, value = covariates, colSpan = 6, rowSpan = 3, fontSize = 15)
xlsx.addTable(wb, sheet, PFAS_cpc_lm)
xlsx.addLineBreak(sheet, 1)
xlsx.addParagraph(wb, sheet, value = "mixture analysis", colSpan = 4, rowSpan = 2, fontSize = 15)
xlsx.addTable(wb, sheet, GAA_mixture)
#saveWorkbook(wb, filename)
#xlsx.openFile(filename)
```


#GAA ~ PFAS * sex (for loop )
```{r}
PFAS_cpc_int = data.frame()

#pheno = pheno %>% filter(.$CPC_lm_residuals > -0.01)
for (i in c("PFOA","PFHxS","PFNA","PFDA","PFOS")) {
  model = lm(paste0("CPC_lm_residuals~", i,"+ mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category + childs_sex:",i), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[10,4]
  estimate <- coef(model)[10]
  df = data.frame(p.value = p_value, estimate = estimate,  PFAS = i)
  PFAS_cpc_int = rbind(PFAS_cpc_int, df)
}

PFAS_cpc_int
```

#GAA ~ PFAS mixture stratified by sex 
```{r}
print("males")
m = qgcomp(CPC_lm_residuals ~  PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category + gest_age_in_weeks_edd , expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno[which(pheno$childs_sex == "Male"),], q = 4)


print("Female")
f = qgcomp(CPC_lm_residuals ~  PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category + gest_age_in_weeks_edd, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno[which(pheno$childs_sex == "Female"),], q = 4)

mixture_stratified = data.frame(sex = c("male", "female"), estimate =  c(m[["psi"]], f[["psi"]]), p.value = c(m[["pval"]][2], f[["pval"]][2]))
```

#GAA ~ PFAS stratified for females (independent PFAS analysis )
```{r}
PFAS_cpc_lm_f = data.frame()

#pheno = pheno %>% filter(.$CPC_lm_residuals > -0.01)
for (i in c("PFOA","PFHxS","PFNA","PFDA","PFOS")) {
  model = lm(paste0("CPC_lm_residuals~", i,"+ mom_age_at_birth + mom_education + MomBMI_36wks +  MomBMI_10wks + official_enroll_category "), data = pheno[which(pheno$childs_sex == "Female"),])
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate,  PFAS = i)
  PFAS_cpc_lm_f = rbind(PFAS_cpc_lm_f, df)
}

PFAS_cpc_lm_f
```

#GAA ~ PFAS stratified for male (independent PFAS analysis )
```{r}
PFAS_cpc_lm_m = data.frame()

#pheno = pheno %>% filter(.$CPC_lm_residuals > -0.01)
for (i in c("PFOA","PFHxS","PFNA","PFDA","PFOS")) {
  model = lm(paste0("CPC_lm_residuals~", i,"+ mom_age_at_birth + mom_education + MomBMI_36wks +  MomBMI_10wks + official_enroll_category"), data = pheno[which(pheno$childs_sex == "Male"),])
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate,  PFAS = i)
  PFAS_cpc_lm_m = rbind(PFAS_cpc_lm_m, df)
}

PFAS_cpc_lm_m
```

#mixture interaction analysis 
```{r}
library(qgcompint)
qfitemm <- qgcomp.emm.noboot(f= CPC_lm_residuals ~ PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category,
           emmvar="childs_sex", 
           expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), 
           data=pheno, q=4, 
           family=gaussian())
mixture_int =  data.frame(estimate =  qfitemm[["psiint"]], p.value = qfitemm[["pval"]][4])
```


##sheet 2 of xlsx
```{r}
sheet2 = createSheet(wb, "Sheet 2")
xlsx.addHeader(wb, sheet2, value = "Biological Sex Analysis", level = 1, color = "black", underline = 1)
xlsx.addLineBreak(sheet, 1)
covariates = "covariates: mom_age_at_birth, mom_education, MomBMI_36wks, childs_sex, MomBMI_10wks, official_enroll_category"
xlsx.addParagraph(wb, sheet2, value = covariates, colSpan = 6, rowSpan = 3, fontSize = 15)
xlsx.addLineBreak(sheet2, 1)
xlsx.addHeader(wb, sheet2, value = "Sex Interaction", level = 2, color = "black")
xlsx.addTable(wb, sheet2, PFAS_cpc_int)
xlsx.addHeader(wb, sheet2, value = "Straitifed Analysis", level = 2, color = "black")
xlsx.addParagraph(wb, sheet2, value = "Female", colSpan = 1, rowSpan = 1, fontSize = 13, isBold = T)
xlsx.addTable(wb, sheet2, PFAS_cpc_lm_f)
xlsx.addParagraph(wb, sheet2, value = "Male", colSpan = 1, rowSpan = 1, fontSize = 13, isBold = T)
xlsx.addTable(wb, sheet2, PFAS_cpc_lm_m)
xlsx.addHeader(wb, sheet2, value = "Straitifed Mixture Analysis", level = 2, color = "black")
xlsx.addTable(wb, sheet2, mixture_stratified)
xlsx.addHeader(wb, sheet2, value = "Mixture Analysis with Sex Interaction", level = 2, color = "black")
xlsx.addTable(wb, sheet2, mixture_int)
#saveWorkbook(wb, filename)
#xlsx.openFile(filename)
```


#cell type proportion ~ PFAS mixture
```{r}

t = qgcomp(Trophoblasts ~  PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category + gest_age_in_weeks_edd , expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)


s = qgcomp(Stromal ~  PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category + gest_age_in_weeks_edd , expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)


h = qgcomp(Hofbauer ~  PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education +  MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category + gest_age_in_weeks_edd, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)


e = qgcomp(Endothelial ~ PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education +  MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category + gest_age_in_weeks_edd, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)


n = qgcomp(nRBC ~ PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education +  MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category + gest_age_in_weeks_edd, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)


st = qgcomp(Syncytiotrophoblast ~  PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education +  MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category + gest_age_in_weeks_edd, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)

mixture_cells = data.frame(cell_type = c("Trophoblasts", "Stromal", "Hofbauer", "Endothelial","nRBC", "Syncytiotrophoblast"), estimate =  c(t[["psi"]], s[["psi"]], h[["psi"]], e[["psi"]], n[["psi"]], st[["psi"]]), p.value = c(t[["pval"]][2], s[["pval"]][2], h[["pval"]][2], e[["pval"]][2], n[["pval"]][2], st[["pval"]][2]))
```




```{r}
outcome_list <- c("Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")

mixture_cells_m = data.frame()
   for ( i in 1:length(outcome_list)){
    mod1.formula <- as.formula(paste0(outcome_list[i], "~official_enroll_category", 
                                       "+mom_age_at_birth",
                                       "+gest_age_in_weeks_edd",
                                       "+mom_education", # cell mix  
                                       "+PFOS",  "+PFNA", "+PFHxS",  "+PFDA", "+PFOA"))
      q_gcomp <- qgcomp::qgcomp.noboot(mod1.formula,                                         # pfas 
                                     expnms= c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"),
                                     dat = pheno[which(pheno$childs_sex == "Male"),], 
                                     family=gaussian(), 
                                     q=4) 
      
       df = data.frame(cell = outcome_list[i], estimate = q_gcomp[["psi"]], p.value = q_gcomp[["pval"]][2])
  mixture_cells_m = rbind(mixture_cells_m, df)
   }
  


mixture_cells_f = data.frame()
   for ( i in 1:length(outcome_list)){
    mod1.formula <- as.formula(paste0(outcome_list[i], "~official_enroll_category", 
                                       "+mom_age_at_birth",
                                       "+gest_age_in_weeks_edd",
                                       "+mom_education", # cell mix  
                                       "+PFOS",  "+PFNA", "+PFHxS",  "+PFDA", "+PFOA"))
      q_gcomp <- qgcomp::qgcomp.noboot(mod1.formula,                                         # pfas 
                                     expnms= c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"),
                                     dat = pheno[which(pheno$childs_sex == "Female"),], 
                                     family=gaussian(), 
                                     q=4) 
      
       df = data.frame(cell = outcome_list[i], estimate = q_gcomp[["psi"]], p.value = q_gcomp[["pval"]][2])
  mixture_cells_f = rbind(mixture_cells_f, df)
   }
```

##Sheet 3 of xlsx
```{r}
sheet3 = createSheet(wb, "Sheet3")
xlsx.addHeader(wb, sheet3, value = "Cell Type Proportions~PFAS mixture", level = 2, color = "black")
xlsx.addTable(wb, sheet3, mixture_cells)
xlsx.addHeader(wb, sheet3, value = "Sex Stratified Cell Type Proportions~PFAS mixture", level = 2, color = "black")
xlsx.addParagraph(wb, sheet3, value = "Male", colSpan = 1, rowSpan = 1, fontSize = 13, isBold = T)
xlsx.addTable(wb, sheet3, mixture_cells_m)
xlsx.addParagraph(wb, sheet3, value = "Female", colSpan = 1, rowSpan = 1, fontSize = 13, isBold = T)
xlsx.addTable(wb, sheet3, mixture_cells_f)
#saveWorkbook(wb, filename)
#xlsx.openFile(filename)
```

#creating a data frame of  stratified mixture results of each outcome variable 
```{r}
#"Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast"
results = data.frame()
outcome_list <- c("Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")
 for(sex in c("Male", "Female")){
   for ( i in 1:length(outcome_list)){
    mod1.formula <- as.formula(paste0(outcome_list[i], "~official_enroll_category", 
                                       "+mom_age_at_birth",
                                       "+gest_age_in_weeks_edd",
                                       "+mom_education", # cell mix  
                                       "+PFOS",  "+PFNA", "+PFHxS",  "+PFDA", "+PFOA"))
      q_gcomp <- qgcomp::qgcomp.noboot(mod1.formula,                                         # pfas 
                                     expnms= c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"),
                                     dat = pheno[which(pheno$childs_sex == sex),], 
                                     family=gaussian(), 
                                     q=4) 
    
    
    # Output results from mixture 
    psi<-round(q_gcomp$psi,4) # extract overall mixture estimate
    # pvals and adjust it
    pval <- q_gcomp[["pval"]][[2]]
    # 95%ci
    psi.ci.ll<-round(q_gcomp$ci.coef,5)[2] #extract CI for mixture
    psi.ci.ul<-round(q_gcomp$ci.coef,5)[4] #extract CI for mixture
    
    # figure 1.  heatmap for plot(qgcomp)
    # use weights from
    weights = c(q_gcomp[["pos.weights"]], -q_gcomp[["neg.weights"]])
    weights = data.frame(t(weights))
    weights <- weights[, c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS")]
    
    # figure 2. heatmap
    pos.psi<-round(q_gcomp$pos.psi,5) #extract sum of positive weights for mixture
    neg.psi<-round(q_gcomp$neg.psi,5) #extract sum of negative weights for mixture
    
    
    df = data.frame( dependent_variable = outcome_list[i], 
                      psi = psi, 
                      psi.ci.ll = psi.ci.ll, 
                      psi.ci.ul = psi.ci.ul, 
                      pval = pval,
                      pos.psi=  pos.psi,
                      neg.psi=  neg.psi,
                      PFOS =  weights$PFOS,
                      PFNA =  weights$PFHxS,
                      PFHxS =  weights$PFHxS,
                      PFDA =   weights$PFDA,
                      PFOA=  weights$PFOA,
                      sex = sex)
    results = rbind(results, df)
}

}

```

#using data from above to create heat map 
```{r}
plot_qgcomp_pca_heatmap <- gather(results, key="pfas", value="weights", 8:12)
plot_qgcomp_pca_heatmap$weights <- round(plot_qgcomp_pca_heatmap$weights, 2)

myPalette <- colorRampPalette(rev(RColorBrewer::brewer.pal(11, "PuOr")))


ggplot(plot_qgcomp_pca_heatmap , aes( 
  #y = reorder(paste0(cpgs), psi),
  y = reorder(dependent_variable, psi), # reorder cpgs based on size of effect estimates, and add chr, gene and annotoations to y axis as labels
  fill = weights, color = weights, 
  x = pfas)) +
  geom_tile() +
  scale_colour_gradientn(colours = myPalette(100)) +
  scale_fill_gradientn(colours = myPalette(100)) +
  labs(y = "Dependent Variable", 
       x = "", fill = "Weights", color= "Weights") +
  theme_bw() +
    facet_grid(~ sex, scales = "free_y") +
 
  theme(panel.border = element_blank(), 
        legend.position="top",
        axis.text.x = element_text(angle= 0), 
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size=11, face="bold", colour = "black"),
        axis.title.y = element_text(size=11, face="bold", colour = "black"))


```
#forest plot of data frame of stratified analysis 
```{r}
library(forestplot)
ggplot(plot_qgcomp_pca_heatmap,
                        aes(x = psi, xmin = psi.ci.ll, xmax = psi.ci.ul,
                            y = dependent_variable)) +
  geom_vline(xintercept = 0, linetype = "longdash", position=position_dodge(width = 0.5)) +
  geom_errorbarh(height = 0.2, lwd=1.05, position=position_dodge(width = 0.5)) +
  geom_point(size = 2, shape = "circle", stroke = 0.5, position=position_dodge(width = 0.5)) +
  # xlim(c(-2, 2))+
  # geom_text(aes(y=as.double(Beta), label=as.double(Beta)), 
  #           position = position_dodge2(width=4), vjust = -10) +
  #  Un-comment the above line to check the effect estimates. The colors and effect estimates should match up with the tab_model() estimates from above.
  xlab("Effect Estimate (95% CI)") +
  ylab(" ")  +
  facet_wrap(~sex) + #strip.position="left", nrow=72, scales = "free_y")+
  theme_bw() +theme(panel.border = element_blank(), legend.position="top", 
                    axis.title.x = element_text(size=9, face="bold", colour = "black"), 
                    axis.text.y = element_text(size=9, face="bold", colour = "black")) #  For bold axis, include the following in theme(): axis.title.x = element_text(size=9, face="bold", colour = "black"), axis.text.y = element_text(size=9, face="bold", colour = "black")

```


#GAA ~ cell type proportion adjusted for covariates
```{r}
GAA_df = data.frame()
for( cell in c( "Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")){
   model = lm(paste0("CPC_lm_residuals~", cell, "+ mom_age_at_birth + mom_education + MomBMI_36wks +  MomBMI_10wks + official_enroll_category + childs_sex"), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate, cell = cell)
  GAA_df = rbind(GAA_df, df)
}

GAA_df
```

#GA ~ cell type proportion adjusted for covariates 
```{r}
GA_df = data.frame()
for( cell in c( "Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")){
   model = lm(paste0("gest_age_in_weeks_edd~", cell, "+ mom_age_at_birth + mom_education + MomBMI_36wks +  MomBMI_10wks + official_enroll_category + childs_sex"), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate, cell = cell)
  GA_df = rbind(GA_df, df)
}

GA_df
```
##Sheet 4 of xlsx
```{r}
sheet4 = createSheet(wb, "Sheet4")
xlsx.addHeader(wb, sheet4, value = "GAA ~ Individual Cell Type Proportions ", level = 2, color = "black")
xlsx.addParagraph(wb, sheet4, value = covariates, colSpan = 6, rowSpan = 3, fontSize = 13, isBold = T)
xlsx.addTable(wb, sheet4, GAA_df)
xlsx.addHeader(wb, sheet4, value = "GA ~ Individual Cell Type Proportions", level = 2, color = "black")
xlsx.addTable(wb, sheet4, GA_df)
#saveWorkbook(wb, filename)
#xlsx.openFile(filename)
```


```{r}
PFAS_bw_lm = data.frame()

#pheno = pheno %>% filter(.$CPC_lm_residuals > -0.01)
for (i in c("PFOA","PFHxS","PFNA","PFDA","PFOS")) {
  model = lm(paste0("birth_wt_kg~", i,"+ mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category "), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate,  PFAS = i)
  PFAS_bw_lm = rbind(PFAS_bw_lm, df)
}

PFAS_bw_lm
```

```{r}
PFAS_bl_lm = data.frame()

#pheno = pheno %>% filter(.$CPC_lm_residuals > -0.01)
for (i in c("PFOA","PFHxS","PFNA","PFDA","PFOS")) {
  model = lm(paste0("birth_len_cm~", i,"+ mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category "), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate,  PFAS = i)
  PFAS_bl_lm = rbind(PFAS_bl_lm, df)
}

PFAS_bl_lm
```

```{r}
mixture_BL = qgcomp( birth_len_cm~ PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)

mixture_BL = data.frame(estimate =  mixture_BL[["psi"]], p.value = mixture_BL[["pval"]][2])
```

```{r}
mixture_BW = qgcomp(birth_wt_kg~ PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)

mixture_BW = data.frame(estimate =  mixture_BW[["psi"]], p.value = mixture_BW[["pval"]][2])
```
##Birthlength and each cell type 
**Birth length was not associated with any cell type **
```{r}
BL_cell_df = data.frame()
for( cell in c( "Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")){
   model = lm(paste0("birth_len_cm~", cell, "+ mom_age_at_birth + mom_education + MomBMI_36wks +  MomBMI_10wks + official_enroll_category + childs_sex"), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate, cell = cell)
 BL_cell_df = rbind(BL_cell_df, df)
}

BL_cell_df
```
##Birth weight and each cell type 
**Birth weight was not associated with any cell type **
```{r}
BW_cell_df = data.frame()
for( cell in c( "Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")){
   model = lm(paste0("birth_wt_kg~", cell, "+ mom_age_at_birth + mom_education + MomBMI_36wks +  MomBMI_10wks + official_enroll_category + childs_sex"), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate, cell = cell)
 BW_cell_df = rbind(BW_cell_df, df)
}

BW_cell_df
```


#Sheet 5 of xlsx 
```{r}
sheet5 = createSheet(wb, "Sheet5")
xlsx.addHeader(wb, sheet5, value = "Birth Outcomes and PFAS ", level = 1, color = "black", underline = 2)
xlsx.addHeader(wb, sheet5, value = "Birth Weight~ Individual PFAS", level = 2, color = "black")
xlsx.addParagraph(wb, sheet5, value = covariates, colSpan = 6, rowSpan = 3, fontSize = 13, isBold = T)
xlsx.addTable(wb, sheet5, PFAS_bw_lm )
xlsx.addHeader(wb, sheet5, value = "Birth Length~ Individual PFAS", level = 2, color = "black")
xlsx.addTable(wb, sheet5, GA_df)
xlsx.addHeader(wb, sheet5, value = "Birth Weight and PFAS Mixture", level = 2, color = "black")
xlsx.addTable(wb, sheet5, mixture_BW )
xlsx.addHeader(wb, sheet5, value = "Birth Length and PFAS Mixture", level = 2, color = "black")
xlsx.addTable(wb, sheet5, mixture_BL )
#saveWorkbook(wb, filename)
#xlsx.openFile(filename)
```


#Sheet 6 of xlsx 
```{r}
sheet6 = createSheet(wb, "Sheet6")
xlsx.addHeader(wb, sheet6, value = "Birth Outcomes and GAA ", level = 1, color = "black", underline = 2)
xlsx.addParagraph(wb, sheet6, value = covariates, colSpan = 6, rowSpan = 3, fontSize = 13, isBold = T)
xlsx.addHeader(wb, sheet6, value = "Birth Weight~ GAA", level = 2, color = "black")
xlsx.addTable(wb, sheet6, BW_GAA)
xlsx.addHeader(wb, sheet6, value = "Birth Length ~ GAA", level = 2, color = "black")
xlsx.addTable(wb, sheet6, BL_GAA )
xlsx.addHeader(wb, sheet6, value = "Birth Length~GAA Sex Stratified and Interaction", level = 2, color = "black")
xlsx.addTable(wb, sheet6, result_table)
xlsx.addHeader(wb, sheet6, value = "Birth Weight~GAA Sex Stratified and Interaction", level = 2, color = "black")
xlsx.addTable(wb, sheet6, BW)
#saveWorkbook(wb, filename)
#xlsx.openFile(filename)
```

```{r}
BW_GAA <- data.frame(
  Dependent_Variable = c("GA", "GAA"),
  Estimate = NA,
  P_Value = NA
)

# Model for GA
model_ga <- lm(birth_wt_kg ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno)
summary_ga <- summary(model_ga)
BW_GAA$Estimate[1] <- coef(summary_ga)["CPC_lm_residuals", "Estimate"]
BW_GAA$P_Value[1] <- coef(summary_ga)["CPC_lm_residuals", "Pr(>|t|)"]

# Model for GAA
model_gaa <- lm(birth_wt_kg ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno)
summary_gaa <- summary(model_gaa)
BW_GAA$Estimate[2] <- coef(summary_gaa)["CPC_lm_residuals", "Estimate"]
BW_GAA$P_Value[2] <- coef(summary_gaa)["CPC_lm_residuals", "Pr(>|t|)"]

# Print the result table
print(BW_GAA)
```

```{r}
# Create an empty data frame to store results
BL_GAA <- data.frame(
  Dependent_Variable = c("GA", "GAA"),
  Estimate = NA,
  P_Value = NA
)

# Model for GA
model_ga <- lm(birth_len_cm ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno)
summary_ga <- summary(model_ga)
BL_GAA$Estimate[1] <- coef(summary_ga)["CPC_lm_residuals", "Estimate"]
BL_GAA$P_Value[1] <- coef(summary_ga)["CPC_lm_residuals", "Pr(>|t|)"]

# Model for GAA
model_gaa <- lm(birth_len_cm ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno)
summary_gaa <- summary(model_gaa)
BL_GAA$Estimate[2] <- coef(summary_gaa)["CPC_lm_residuals", "Estimate"]
BL_GAA$P_Value[2] <- coef(summary_gaa)["CPC_lm_residuals", "Pr(>|t|)"]

# Print the result table
print(BL_GAA)

```


```{r}
# Create an empty data frame to store results
result_table <- data.frame(
  Row = c("Male", "Female", "Interaction"),
  Estimate = NA,
  P_Value = NA
)

# Model for Male
model_male <- lm(birth_len_cm ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno[which(pheno$childs_sex == "Male"),])
summary_male <- summary(model_male)
result_table$Estimate[1] <- coef(summary_male)["CPC_lm_residuals", "Estimate"]
result_table$P_Value[1] <- coef(summary_male)["CPC_lm_residuals", "Pr(>|t|)"]

# Model for Female
model_female <- lm(birth_len_cm ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno[which(pheno$childs_sex == "Female"),])
summary_female <- summary(model_female)
result_table$Estimate[2] <- coef(summary_female)["CPC_lm_residuals", "Estimate"]
result_table$P_Value[2] <- coef(summary_female)["CPC_lm_residuals", "Pr(>|t|)"]

# Model for Interaction
model_interaction <- lm(birth_len_cm ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category + childs_sex + CPC_lm_residuals:childs_sex, data = pheno)
summary_interaction <- summary(model_interaction)
result_table$Estimate[3] <- summary_interaction$coefficients["CPC_lm_residuals:childs_sexMale","Estimate"]
result_table$P_Value[3] <- coef(summary_interaction)["CPC_lm_residuals:childs_sexMale", "Pr(>|t|)"]

# Print the result table
print(result_table)

```


```{r}
# Create an empty data frame to store results
BW <- data.frame(
  Row = c("Male", "Female", "Interaction"),
  Estimate = NA,
  P_Value = NA
)

# Model for Male
model_male <- lm(birth_wt_kg~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno[which(pheno$childs_sex == "Male"),])
summary_male <- summary(model_male)
BW$Estimate[1] <- coef(summary_male)["CPC_lm_residuals", "Estimate"]
BW$P_Value[1] <- coef(summary_male)["CPC_lm_residuals", "Pr(>|t|)"]

# Model for Female
model_female <- lm(birth_wt_kg ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno[which(pheno$childs_sex == "Female"),])
summary_female <- summary(model_female)
BW$Estimate[2] <- coef(summary_female)["CPC_lm_residuals", "Estimate"]
BW$P_Value[2] <- coef(summary_female)["CPC_lm_residuals", "Pr(>|t|)"]

# Model for Interaction
model_interaction <- lm(birth_wt_kg~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category + childs_sex + CPC_lm_residuals:childs_sex, data = pheno)
summary_interaction <- summary(model_interaction)
BW$Estimate[3] <- summary_interaction$coefficients["CPC_lm_residuals:childs_sexMale","Estimate"]
BW$P_Value[3] <- coef(summary_interaction)["CPC_lm_residuals:childs_sexMale", "Pr(>|t|)"]

# Print the result table
print(BW)

```


```{r}
library(zscorer)
library(tidyverse)
zscorer_dat <- tibble(
  id = pheno$participant_id,
  sex = ifelse(pheno$childs_sex == "Male",1, 2),
   weight_kg = pheno$birth_wt_kg,
  length_cm = pheno$birth_len_cm,
  age_days = pheno$gest_age_in_weeks_edd # can calculate from DOB and the date of the visit
)

zscorer_calcs <- zscorer_dat %>%
 
  # Add WFL z-score
  zscorer::addWGSR(
    data = .,
    sex = "sex",
    firstPart = "weight_kg",
    secondPart = "length_cm",
    index = "wfl"
  ) %>%
  
  # Add WFA z-score
  zscorer::addWGSR(
    data = .,
    sex = "sex",
    firstPart = "weight_kg", # change to "length_cm" for "lfa"
    secondPart = "age_days",
    index = "wfa" # "lfa" for length
  )

pheno$wflz = zscorer_calcs$wflz

pheno$wfaz = zscorer_calcs$wfaz
```

#Sheet 7 of xlsx 
```{r}
sheet7 = createSheet(wb, "Sheet7")
xlsx.addHeader(wb, sheet7, value = "Z score of Birth Weight and Birth Length ", level = 1, color = "black", underline = 2)
xlsx.addParagraph(wb, sheet7, value = covariates, colSpan = 6, rowSpan = 3, fontSize = 13, isBold = T)
xlsx.addParagraph(wb, sheet7, value = "weight for length (wfl) and weight for age (wfa)", colSpan = 5, rowSpan = 5, fontSize = 13, isBold = T)
xlsx.addHeader(wb, sheet7, value = "wfl/wfa ~ GAA", level = 2, color = "black")
xlsx.addTable(wb, sheet7, wfl_GAA )
xlsx.addHeader(wb, sheet7, value = "wfaz~GAA Sex Stratified and Interaction", level = 2, color = "black")
xlsx.addTable(wb, sheet7, wfaz)
xlsx.addHeader(wb, sheet7, value = "wflz~GAA Sex Stratified and Interaction", level = 2, color = "black")
xlsx.addTable(wb, sheet7, wflz)
xlsx.addHeader(wb, sheet7, value = "wflz/wfaz~PFAS mixture", level = 2, color = "black")
xlsx.addTable(wb, sheet7, mixture)
xlsx.addHeader(wb, sheet7, value = "wflz~ individual PFAS", level = 2, color = "black")
xlsx.addTable(wb, sheet7, PFAS_wfl_lm)
xlsx.addHeader(wb, sheet7, value = "wfaz~ individual PFAS", level = 2, color = "black")
xlsx.addTable(wb, sheet7, PFAS_wfa_lm)
saveWorkbook(wb, filename)
xlsx.openFile(filename)
```

```{r}

# Create an empty data frame to store results
 wfl_GAA <- data.frame(
  Independent_Variable = c("wflz", "wfaz"),
  Estimate = NA,
  P_Value = NA
)

# Model for GA
model_ga <- lm(wflz ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno)
summary_ga <- summary(model_ga)
wfl_GAA$Estimate[1] <- coef(summary_ga)["CPC_lm_residuals", "Estimate"]
wfl_GAA$P_Value[1] <- coef(summary_ga)["CPC_lm_residuals", "Pr(>|t|)"]

# Model for GAA
model_gaa <- lm(wfaz ~ CPC_lm_residuals + mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category, data = pheno)
summary_gaa <- summary(model_gaa)
wfl_GAA$Estimate[2] <- coef(summary_gaa)["CPC_lm_residuals", "Estimate"]
wfl_GAA$P_Value[2] <- coef(summary_gaa)["CPC_lm_residuals", "Pr(>|t|)"]

# Print the result table
print(wfl_GAA)

```

```{r}
wfaz <- data.frame(
  Row = c("Male", "Female", "Interaction"),
  Estimate = NA,
  P_Value = NA
)

summary_interaction = summary(lm(wfaz ~ CPC_lm_residuals+ mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category + childs_sex + CPC_lm_residuals:childs_sex, data = pheno))
wfaz$Estimate[3] <- summary_interaction$coefficients["CPC_lm_residuals:childs_sexMale","Estimate"]
wfaz$P_Value[3] <- coef(summary_interaction)["CPC_lm_residuals:childs_sexMale", "Pr(>|t|)"]

summary_male <- summary(lm(wfaz ~ CPC_lm_residuals+ mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category , data = pheno[which(pheno$childs_sex == "Male"), ]))
wfaz$Estimate[1] <- coef(summary_male)["CPC_lm_residuals", "Estimate"]
wfaz$P_Value[1] <- coef(summary_male)["CPC_lm_residuals", "Pr(>|t|)"]

summary_female <- summary(lm(wfaz ~ CPC_lm_residuals+ mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category , data = pheno[which(pheno$childs_sex == "Female"), ]))
wfaz$Estimate[2] <- coef(summary_female)["CPC_lm_residuals", "Estimate"]
wfaz$P_Value[2] <- coef(summary_female)["CPC_lm_residuals", "Pr(>|t|)"]
```

```{r}
wflz <- data.frame(
  Row = c("Male", "Female", "Interaction"),
  Estimate = NA,
  P_Value = NA
)

summary_interaction = summary(lm(wflz ~ CPC_lm_residuals+ mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category + childs_sex + CPC_lm_residuals:childs_sex, data = pheno))
wflz$Estimate[3] <- summary_interaction$coefficients["CPC_lm_residuals:childs_sexMale","Estimate"]
wflz$P_Value[3] <- coef(summary_interaction)["CPC_lm_residuals:childs_sexMale", "Pr(>|t|)"]

summary_male <- summary(lm(wflz ~ CPC_lm_residuals+ mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category , data = pheno[which(pheno$childs_sex == "Male"), ]))
wflz$Estimate[1] <- coef(summary_male)["CPC_lm_residuals", "Estimate"]
wflz$P_Value[1] <- coef(summary_male)["CPC_lm_residuals", "Pr(>|t|)"]

summary_female <- summary(lm(wflz~ CPC_lm_residuals+ mom_age_at_birth + mom_education + MomBMI_36wks + MomBMI_10wks + official_enroll_category , data = pheno[which(pheno$childs_sex == "Female"), ]))
wflz$Estimate[2] <- coef(summary_female)["CPC_lm_residuals", "Estimate"]
wflz$P_Value[2] <- coef(summary_female)["CPC_lm_residuals", "Pr(>|t|)"]
```

```{r}
 
mixture_wfl= qgcomp( wflz ~ PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)


mixture_wfa= qgcomp( wfaz ~ PFOA + PFHxS + PFNA + PFDA + PFOS + mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category, expnms = c("PFOA","PFHxS", "PFNA", "PFDA", "PFOS"), data = pheno, q = 4)

mixture = data.frame(dependent_variable = c("wflz", "wfaz"), estimate =  c(mixture_wfl[["psi"]], mixture_wfa[["psi"]]), p.value = c(mixture_wfl[["pval"]][2], mixture_wfa[["pval"]][2]))
```

```{r}
PFAS_wfl_lm = data.frame()

#pheno = pheno %>% filter(.$CPC_lm_residuals > -0.01)
for (i in c("PFOA","PFHxS","PFNA","PFDA","PFOS")) {
  model = lm(paste0("wflz~", i,"+ mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category "), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate,  PFAS = i)
  PFAS_wfl_lm = rbind(PFAS_wfl_lm, df)
}

PFAS_wfl_lm
```

```{r}
PFAS_wfa_lm = data.frame()

#pheno = pheno %>% filter(.$CPC_lm_residuals > -0.01)
for (i in c("PFOA","PFHxS","PFNA","PFDA","PFOS")) {
  model = lm(paste0("wfaz~", i,"+ mom_age_at_birth + mom_education + MomBMI_36wks +  childs_sex + MomBMI_10wks + official_enroll_category "), data = pheno)
  lm_summary <- summary(model)
  p_value <- lm_summary$coefficients[2,4]
  estimate <- coef(model)[2]
  df = data.frame(p.value = p_value, estimate = estimate,  PFAS = i)
  PFAS_wfa_lm = rbind(PFAS_wfa_lm, df)
}

PFAS_wfa_lm
```