```{r}
load("/Users/cynthiaperez/GitHub/Placenta_RNAseq_WGCNA/april24_input.Rdata")
library(planet)
library(tidyverse)
```


```{r}
load("/Users/cynthiaperez/GitHub/Placenta_RNAseq_WGCNA/april24_input.Rdata")
load("2_BMIQ_Adjsuted.RData")
load("ALLBETAS.FunNormBMIQ.RData")
colnames(FunNorm.BMIQ.ALL) = pd$Sample_Name

FunNorm.BMIQ.ALL = FunNorm.BMIQ.ALL[,row.names(pfas_df_cp)]
pheno = pheno[row.names(pfas_df_cp),]
pheno$RPC_clock = predictAge(FunNorm.BMIQ.ALL, type = "RPC")
pheno$CPC_clock = predictAge(FunNorm.BMIQ.ALL, type = "CPC")
pheno$RRPC_clock = predictAge(FunNorm.BMIQ.ALL, type = "RRPC")

pheno$RPC_lm_residuals = summary(lm(RPC_clock~gest_age_in_weeks_edd, pheno))$residuals
pheno$RRPC_lm_residuals = summary(lm(RRPC_clock~gest_age_in_weeks_edd, pheno))$residuals
pheno$CPC_lm_residuals= summary(lm(CPC_clock~gest_age_in_weeks_edd, pheno))$residuals

cor.test(pheno$gest_age_in_weeks_edd, pheno$CPC_clock)
cor.test(pheno$gest_age_in_weeks_edd, pheno$RPC_clock)
cor.test(pheno$gest_age_in_weeks_edd, pheno$RRPC_clock)

```

covariates to adjust for maternal age at birth, maternal education level, pre-pregnancy BMI, gestational weight gain, and infant sex


the phenotype profiles I have contain different information 
```{r}

PFASp = c("L-PFOA","PFHxS","PFNA","PFDA","s-PFOS","sum1", "sum2" )
pfas_log2 = log2(pfas_df_cp[, PFASp])
pfas_log2$ID = row.names(pfas_log2)

colnames(pheno)[1] = "ID"
merged_df <- left_join(pheno, pfas_log2, by = "ID")
colnames(merged_df)[60] = "PFOS"
colnames(merged_df)[56] = "PFOA"

PFASp = c("PFOA","PFHxS","PFNA","PFDA","PFOS","sum1", "sum2" )

PFAS_rpc_lm = data.frame()


for (i in seq_along(PFASp)) {
  lm = summary(lm(paste0("RPC_lm_residuals~", PFASp[i],"+ Trophoblasts + Stromal + Hofbauer + Endothelial + nRBC + Syncytiotrophoblast + mom_education + MomBMI_10wks + MomBMI_36wks + child_sex + gest_age_in_weeks_edd + mom_age_at_birth"), data = merged_df))$coefficients[2,4] 
   df = data.frame(p.value = lm, PFAS = PFASp[i])
    PFAS_rpc_lm = rbind(PFAS_rpc_lm, df)
}

PFAS_rpc_lm
```

```{r}
PFAS_cpc_lm = data.frame()


for (i in seq_along(PFASp)) {
  lm = summary(lm(paste0("CPC_lm_residuals~", PFASp[i],"+ Trophoblasts + Stromal + Hofbauer + Endothelial + nRBC + Syncytiotrophoblast + mom_education + MomBMI_10wks + MomBMI_36wks + child_sex + gest_age_in_weeks_edd + mom_age_at_birth"), data = merged_df))
   df = data.frame(p.value = lm$coefficients[2,4] , PFAS = PFASp[i], estimate = lm$coefficients[2,1])
    PFAS_cpc_lm = rbind(PFAS_cpc_lm, df)
}

PFAS_cpc_lm
```

#run a model where there are no adjustments 
add everything to a word doc for epigenetic clocks and wgcna 
```{r}
PFASp = c("PFOA","PFHxS","PFNA","PFDA","PFOS","sum1", "sum2" )

PFAS_rpc_lm_no_cov = data.frame()


for (i in seq_along(PFASp)) {
  lm = summary(lm(paste0("RPC_lm_residuals~", PFASp[i]), data = merged_df))$coefficients[2,4] 
   df = data.frame(p.value = lm, PFAS = PFASp[i])
    PFAS_rpc_lm_no_cov = rbind(PFAS_rpc_lm, df)
}

PFAS_rpc_lm_no_cov
```
```{r}
PFAS_cpc_lm_no_cov = data.frame()


for (i in seq_along(PFASp)) {
  lm = summary(lm(paste0("CPC_lm_residuals~", PFASp[i]), data = merged_df))
   df = data.frame(p.value = lm$coefficients[2,4] , PFAS = PFASp[i], estimate = lm$coefficients[2,1])
    PFAS_cpc_lm_no_cov= rbind(PFAS_cpc_lm, df)
}

PFAS_cpc_lm_no_cov
```
in quartiles PFAS concentrations 
accelerated vs deaccerlated age 


