---
title: "mixture_analysis"
author: "Cynthia Perez"
date: "2025-06-18"
output: html_document
---

## Load Required Libraries

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(qgcomp)
library(qgcompint)
library(ggpubr)
library(car)
library(stringr) 
library(gWQS)
library(dplyr)
```

## Load Data

```{r load-data, message=FALSE}
# Load normalized beta values
load("2_BMIQ_Adjsuted.RData")
load("ALLBETAS.FunNormBMIQ.RData")

# Rename columns to match sample names
colnames(FunNorm.BMIQ.ALL) <- pd$Sample_Name

# Load PFAS data and phenotype info
load("log_PFAS_pheno.Rdata")
```

## Data Cleaning and Subsetting

```{r data-cleaning}
# Retain only samples present in both datasets
cols_to_keep <- intersect(colnames(FunNorm.BMIQ.ALL), rownames(log_pfas))
FunNorm.BMIQ.ALL <- FunNorm.BMIQ.ALL[, cols_to_keep]

# Subset phenotype and PFAS data accordingly
pheno <- pheno[cols_to_keep, ]
log_pfas <- log_pfas[cols_to_keep, ]
log_pfas <- log_pfas %>%
  rename_with(~ str_replace_all(.x, "-", "_"))   # "L-PFOA" -> "L_PFOA"

# Append PFAS values to phenotype data
pfas_vars <- c("L_PFOA","PFHxS","PFNA","PFDA","s_PFOS")
pheno <- cbind(pheno, log_pfas[, pfas_vars])
```

## The variable gwg_kg (gestational weight gain) was in a separate dataset. Adding it to pheno dataset. 

```{r adding-gwg_kg}
load("ViCTER_maternal_environment_2024-05-29.RData")
weight =  wide_pfas_modeling %>% 
          filter(id %in% pheno$participant_id) %>%
          select(id, gwg_kg) %>% 
          #mutate(x = replace_na(gwg_kg, mean(weight$gwg_kg, na.rm = TRUE)))
          dplyr::rename(participant_id = id)


pheno= merge(pheno, weight, by = "participant_id")
rm(list = ls()[!ls() %in% c("pheno")])
```


## Notes

- The following files are not publicly available:
  - `2_BMIQ_Adjsuted.RData`
  - `ALLBETAS.FunNormBMIQ.RData`
  - `log_PFAS_pheno.Rdata`
  - `ViCTER_maternal_environment_2024-05-29.RData`
- Raw beta values are available from (GEO): GSE288358 

## Quantile g-computation main effect
** Dependent variable: GAA, Independent Variable: PFAS Mixture, Covariates: mom age at birth, mom education, child/placenta sex, official enrollment category (lean or overweight), gestational age, gestational weight gain **
```{r}
q_gcomp = qgcomp(CPC_GAA ~ L_PFOA + PFHxS + PFNA + PFDA + s_PFOS + mom_age_at_birth + mom_education + childs_sex + official_enroll_category + gest_age_in_weeks_edd + gwg_kg, expnms = c("L_PFOA","PFHxS", "PFNA", "PFDA", "s_PFOS"), data = pheno, q = 4)

main_CPC = data.frame(variable = "Main Effect", 
                       estimate = q_gcomp[["psi"]], 
                       psi.ci.ll = q_gcomp$ci.coef[2], 
                       psi.ci.ul = q_gcomp$ci.coef[4], 
                       p.value = q_gcomp[["pval"]][2])

main_CPC
```

## quantile g-computation sex stratified outcome CPC GAA 
```{r}

sex_levels <- c("Female", "Male")
out <- vector("list", length(sex))
results_list <- list()

for (sex in sex_levels) {
  dat_sex <- pheno %>% filter(childs_sex == sex)

q_gcomp = qgcomp(CPC_GAA~  L_PFOA + PFHxS + PFNA + PFDA + s_PFOS + mom_age_at_birth + mom_education + official_enroll_category + gest_age_in_weeks_edd + gwg_kg , expnms = c("L_PFOA","PFHxS", "PFNA", "PFDA", "s_PFOS"), data = dat_sex, q = 4)
  

out[[sex]] = data.frame(variable = sex, 
                       estimate = q_gcomp[["psi"]], 
                       psi.ci.ll = q_gcomp$ci.coef[2], 
                       psi.ci.ul = q_gcomp$ci.coef[4], 
                       p.value = q_gcomp[["pval"]][2])
}

CPC_sex_strata <- bind_rows(out)

CPC_sex_strata

```

## quantile g-computation sex and PFAS mixture interaction term outcome CPC_GAA 
```{r}
library(qgcompint)
pheno$childs_sex <- relevel(factor(pheno$childs_sex), ref="Female")
qfitemm <- qgcomp.emm.noboot(f= CPC_GAA ~ L_PFOA + PFHxS + PFNA + PFDA + s_PFOS + mom_age_at_birth + mom_education + gest_age_in_weeks_edd + childs_sex + official_enroll_category + gwg_kg, emmvar="childs_sex", expnms = c("L_PFOA","PFHxS", "PFNA", "PFDA", "s_PFOS"), data=pheno, q=4, family=gaussian())
qfitemm 
getstrateffects(qfitemm , emmval = 1)
GAA_int = data.frame( variable = "ψ x Sex Int.term()",
                      estimate = qfitemm[["psiint"]][["childs_sexMale:mixture"]], 
                       psi.ci.ll = qfitemm$ci.coef[4,1], 
                       psi.ci.ul = qfitemm$ci.coef[4,2], 
                      p.value = qfitemm[["pval"]][4])
```
## quantile g-computation cells types as dependent variable 
```{r}
cells <-  c( "Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")

out <- vector("list", length(cells))
 
results_list <- list()

for (k in cells) {
  mod1.formula <- as.formula(paste0(k, "~official_enroll_category", 
                                       "+mom_age_at_birth",
                                       "+gest_age_in_weeks_edd",
                                       "+mom_education", 
                                       "+gwg_kg",  
                                       "+s_PFOS",  
                                       "+PFNA", 
                                       "+PFHxS", 
                                       "+PFDA", 
                                       "+L_PFOA"))
  

 q_gcomp <- qgcomp::qgcomp.noboot(mod1.formula,                                     
                                     expnms= c("L_PFOA","PFHxS", "PFNA", "PFDA","s_PFOS"),
                                     dat = pheno, 
                                     family=gaussian(), 
                                     q=4) 
  
 out[[k]] = data.frame(variable = "Main Effect", 
                       cell_type = k, 
                       estimate = q_gcomp[["psi"]], 
                       psi.ci.ll = q_gcomp$ci.coef[2], 
                       psi.ci.ul = q_gcomp$ci.coef[4], 
                       p.value = q_gcomp[["pval"]][2])
  
}
results_cells_main_effects <- bind_rows(out)

results_cells_main_effects
```
## quantile g-computation sex stratified cells types as dependent variable 
```{r}
sex_levels <- c("Female", "Male")
cells <- c("Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")

results_list <- list()

for (sex in sex_levels) {
  dat_sex <- pheno %>% filter(childs_sex == sex)
  out <- vector("list", length(cells))
for ( k in cells){
   
  mod1.formula <- as.formula(paste0(k, "~official_enroll_category", 
                                       "+mom_age_at_birth",
                                       "+gest_age_in_weeks_edd",
                                       "+mom_education", 
                                       "+gwg_kg",  
                                       "+s_PFOS",  
                                       "+PFNA", 
                                       "+PFHxS", 
                                       "+PFDA", 
                                       "+L_PFOA"))
  

 q_gcomp <- qgcomp::qgcomp.noboot(mod1.formula,                                     
                                     expnms= c("L_PFOA","PFHxS", "PFNA", "PFDA","s_PFOS"),
                                     dat = dat_sex, 
                                     family=gaussian(), 
                                     q=4) 


  
out[[k]] = data.frame( variable = paste0(sex, " Strata"),  
                       cell_type = k,
                       estimate = q_gcomp[["psi"]], 
                       psi.ci.ll = q_gcomp$ci.coef[2], 
                       psi.ci.ul = q_gcomp$ci.coef[4], 
                       p.value = q_gcomp[["pval"]][2])
} 
  results_list[[sex]] <- bind_rows(out)
}

cells_qgcomp_f = results_list$Female
cells_qgcomp_f
cells_qgcomp_m = results_list$Male
cells_qgcomp_m
```
## quantile g-computation ψ x Sex Int.term() cells types as dependent variable 
```{r}
library(qgcompint)
pheno$childs_sex <- relevel(factor(pheno$childs_sex), ref="Female")

cells <- c("Trophoblasts", "Stromal", "Hofbauer", "Endothelial", "nRBC",  "Syncytiotrophoblast")

results_list <- list()

  out <- vector("list", length(cells))
for ( k in cells){
   
  mod1.formula <- as.formula(paste0(k, "~official_enroll_category", 
                                       "+mom_age_at_birth",
                                       "+gest_age_in_weeks_edd",
                                       "+mom_education", 
                                       "+childs_sex",
                                       "+gwg_kg", 
                                       "+s_PFOS",  
                                       "+PFNA", 
                                       "+PFHxS", 
                                       "+PFDA", 
                                       "+L_PFOA"))
  

 qfitemm <- qgcomp.emm.noboot(f= mod1.formula, 
                              expnms = c("L_PFOA","PFHxS", "PFNA", "PFDA", "s_PFOS"), 
                              emmvar="childs_sex", 
                              data=pheno, 
                              q=4, 
                              family=gaussian())
qfitemm 

getstrateffects(qfitemm , emmval = 1)

out[[k]] = data.frame( variable = "ψ x Sex Int.term()",
                      cell_type = k,
                      estimate = qfitemm[["psiint"]][["childs_sexMale:mixture"]], 
                       psi.ci.ll = qfitemm$ci.coef[4,1], 
                       psi.ci.ul = qfitemm$ci.coef[4,2], 
                      p.value = qfitemm[["pval"]][4])
} 

results_cells_int <- dplyr::bind_rows(out)
results_cells_int 
```

##create csv file - celltypedf 
```{r}
celltypeDF = rbind(results_cells_main_effects, cells_qgcomp_f, cells_qgcomp_m, results_cells_int)
rownames(celltypeDF) = NULL

celltypeDF = celltypeDF %>% arrange(cell_type)
colnames(CellTypeDF) = c("Variable", "Cell.Type", "Estimate", "Lower.CI", "Upper.CI","p.value" )

write.csv(celltypeDF, file = "CellTypeDF.csv", quote = F, row.names = F)
```

##create csv file - qgcomp_gaa
```{r}
qgcomp_gaa = rbind(main_CPC, CPC_sex_strata, GAA_int) 
colnames(qgcomp_gaa) = c("Variable",  "Estimate", "Lower.CI", "Upper.CI", "p.value")
rownames(qgcomp_gaa) = NULL
write.csv(qgcomp_gaa, file = "qgcomp_gaa.csv", quote = F, row.names = F)
```


##forest plot of cell types outcomes 
```{r}
CellTypeDF <- read.csv("CellTypeDF.csv") #made this document 
CellTypeDF$Variable <- factor(CellTypeDF$Variable, levels = c("Main Effect ", "ψ x Sex Int.term()", "Female Strata",  "Male Strata"))
library(forestplot)
ggplot(CellTypeDF, aes(x = Estimate, xmin = Lower.CI, xmax = Upper.CI, y = Cell.Type)) +
  geom_vline(xintercept = 0, linetype = "longdash", position=position_dodge(width = 0.5)) +
  geom_errorbarh(height = 0.2, lwd=1.05, position=position_dodge(width = 0.5)) +
  geom_point(size = 2, shape = "circle", stroke = 0.5, position=position_dodge(width = 0.5)) +
  xlab("Effect Estimate (95% CI)") +
  ylab(" ")  +
  facet_wrap(~Variable) + 
  theme_bw() +theme(panel.border = element_blank(), legend.position="top", 
                    axis.title.x = element_text(size=12, colour = "black"), 
                    axis.text.x = element_text(size=12, colour = "black"), 
                    axis.text.y = element_text(size=12, colour = "black"),
                    strip.text = element_text(size = 12,face="bold")) 
```

##forest plot of cpc GAA outcome 
```{r}
qgcomp_gaa <- read.csv("qgcomp_gaa.csv") #this was 
qgcomp_gaa$Variable <- factor(qgcomp_gaa$Variable, levels = c("Main Effect", "ψ x Sex Int.term()", "Female Strata", "Male Strata"))

library(forestplot)
ggplot(qgcomp_gaa, aes(x = Estimate, xmin = Lower.CI, xmax = Upper.CI, y = Outcome)) +
  geom_vline(xintercept = 0, linetype = "longdash", position=position_dodge(width = 0.5)) +
  geom_errorbarh(height = 0.2, lwd=1.05, position=position_dodge(width = 0.5)) +
  geom_point(size = 2, shape = "circle", stroke = 0.5, position=position_dodge(width = 0.5)) +
  xlab("Effect Estimate (95% CI)") +
  ylab(" ")  +
  facet_wrap(~Variable) + #strip.position="left", nrow=72, scales = "free_y")+
  theme_bw() +theme(panel.border = element_blank(), legend.position="top", 
                    axis.title.x = element_text(size=12, colour = "black"), 
                    axis.text.x = element_text(size=12, colour = "black"), 
                    axis.text.y = element_text(size=12, colour = "black"),
                    strip.text = element_text(size = 12,face="bold")) 
                    # + scale_y_discrete(labels=c("CPC_lm_residuals"="GAA"))
                    
                    #  For bold axis, include the following in theme(): axis.title.x = element_text(size=9, face="bold", colour = "black"), axis.text.y = element_text(size=9, face="bold", colour = "black")
```
## weighted quantile sum preformed on output variable cpc_gaa 
** preformed to confirm qgcomp analysis also preformed for all qgcomp output but all workflow not presented here** 
```{r}
PFASp = c("L_PFOA","PFHxS", "PFNA", "PFDA", "s_PFOS")

results2int <- gwqs(formula = CPC_GAA ~ wqs*childs_sex + mom_age_at_birth + mom_education +  childs_sex + official_enroll_category+ gwg_kg, bint_cat_pos= T, mix_name = PFASp, data = pheno, q = 4, validation = 0, b = 20, plan_strategy = "multicore", family = "gaussian", seed = 10)
summary(results2int)
confint(results2int, level=0.95)

results2int <- gwqs(formula = CPC_GAA ~ wqs*childs_sex + mom_age_at_birth + mom_education +  childs_sex + official_enroll_category+ gwg_kg, bint_cat_pos=FALSE, mix_name = PFASp, data = pheno, q = 4, validation = 0, b = 20, plan_strategy = "multicore", family = "gaussian", seed = 10)
summary(results2int)
confint(results2int, level=0.95)

```
